WED-SQL (Draft)

### 2016-05-12 doc/paper/main_en.tex
paper: intro finalized

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
WED-flow, PostgreSQL, transações longas, ...
\end{IEEEkeywords}

@@

Most of modern business process management software systems are subject to recurrent structural modifications during its
life cycle. This changes are necessary to incorporate new requisites to the software, once it is often impossible 
to come up with all requirements upfront in the modeling phase of project - some of then only appear after the software
is running in production. Furthermore, it is too expensive and time consuming to try to predict all the possibles situations 
and its outcomes, which ultimately may raise exceptions that must be dealt with. Over time, these modifications leads
to code deterioration, exponentially increasing the maintenance costs and compromising the quality of the system as well 
negatively impacting its overall performance.
\\
\indent  Classic business process specification models, for example WS-BPEL, Process Algebra and Petri Networks, focus primarily
on behavioral interactions and relations between process, pushing data analysis into the background and overlooking its
relevance for the project. Therefore, none of these models exactly fits the needs of design dynamic business process 
management systems, which may result in costly system adaptations and maintenance.
\\
\indent  The WED-flow approach, in contrast with the classical models, capture both the inter-process relations and, with the very
same relevance, the data generated by their interactions. With this in mind, the WED-flow allows systems architects to 
think in terms of data states and transitions, similarly to a finite automata designing. Modifying or adding new business 
rules to the project, which is often translated in creation of new processes, is a simple matter of defining new data states 
and conditions for transitions, thus capturing the true nature of dynamic processes and simplifying an incremental evolution 
of the whole project. Real time exceptions are also easier to be handled and integrated to the project, once this approach
is based on transactional properties, thus providing several recovery mechanisms~\cite{ICWS12}.
\\
%problemas da implementacao e justificativa do WED-SQL
%tolerancia a falhas
\indent In order to provide a standard implementation as well as simplifying its software development, in this work we present 
the WED-SQL, an WED-flow framework. The main purpose of WED-SQL is to implement all WED-flow definitions~\cite{FTPM10} and encapsulate 
the control structure, what includes the transactional control, exceptions handling, data states detection, state transitioning 
and ensure data integrity. This framework effectively is an abstraction layer that allow developers to put more effort
on business logic and come up with better WED-flow projects, also making better use of reusable code. 
\\
\indent O WED-SQL foi construído dentro de um SGBD relacional e com isso utiliza a linguagem SQL para definir propriedades
e controles de fluxo de um modelo WED-flow, além de contar com um robusto aparato transacional e ser altamente tolerante
a falhas. Por utilizar como base tecnologias consolidadas e amplamente difundidas no universo da computação, o WED-SQL visa 
disseminar a adoção do paradigma WED-flow na modelagem de processos de negócio , fornecendo uma ferramenta confiável, facilmente
escalável, simples de ser utilizada e que permita a flexibilidade exigida na especificação de controle de fluxo dos processos
de negócio modernos.
\\
\indent O restante desse artigo segue a seguinte estrutura: Na sessão II são apresentados aspectos da arquitetura de software 
e das estruturas de dados internas. A sessão III descreve sucintamente o funcionamento do sistema. A sessão IV descreve
os algoritmos envolvidos. Na sessão V são discutidos em detalhes o gerenciamento de transações e o protocolo de comunicação.
Finalmente, na sessão VI, são apresentados os próximos desafios para expandir as funcionalidades dessa ferramenta.

>>>>

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
WED-flow, Business Process Management, Long Lived Transactions, PostgreSQL, Client-Server Architecture, Distributed Systems
\end{IEEEkeywords}

@@

Most of modern business process management software systems are subject to recurrent structural modifications during its
life cycle. This changes are necessary to incorporate new requisites to the software, once it is often impossible 
to come up with all requirements upfront in the modeling phase of project - some of then only appear after the software
is running in production. Furthermore, it is too expensive and time consuming to try to predict all the possibles situations 
and its outcomes, which ultimately may raise exceptions that must be dealt with. Over time, these modifications leads
to code deterioration, exponentially increasing the maintenance costs and compromising the quality of the system as well 
negatively impacting its overall performance.
\\
\indent  Classic business process specification models, for example WS-BPEL, Process Algebra and Petri Networks, focus primarily
on behavioral interactions and relations between process, pushing data analysis into the background and overlooking its
relevance for the project. Therefore, none of these models exactly fits the needs of design dynamic business process 
management systems, which may result in costly system adaptations and maintenance.
\\
\indent  The WED-flow approach, in contrast with the classical models, capture both the inter-process relations and, with the very
same relevance, the data generated by their interactions. With this in mind, the WED-flow allows systems architects to 
think in terms of data states and transitions, similarly to a finite automata designing. Modifying or adding new business 
rules to the project, which is often translated in creation of new processes, is a simple matter of defining new data states 
and conditions for transitions, thus capturing the true nature of dynamic processes and simplifying an incremental evolution 
of the whole project. Real time exceptions are also easier to be handled and integrated to the project, once this approach
is based on transactional properties, thus providing several recovery mechanisms~\cite{ICWS12}.
\\
%problemas da implementacao e justificativa do WED-SQL
%tolerancia a falhas
\indent In order to provide a standard implementation as well as simplifying its software development, in this work we present 
the WED-SQL, an WED-flow framework. The main purpose of WED-SQL is to implement all WED-flow definitions~\cite{FTPM10} and encapsulate 
the control structure, what includes the transactional control, exceptions handling, data states detection, state transitioning 
and ensure data integrity. This framework effectively is an abstraction layer that allow developers to put more effort
on business logic and come up with better WED-flow projects, also making better use of reusable code. 
\\
\indent The WED-SQL framework has been built inside the PostgreSQL relational database management system, thereby uses SQL language
to specify WED-flow elements, has full transactional support and it is highly fault tolerant. Once it employs well consolidate 
technologies, widespread in the computer science field, this framework targets to disseminate the adoption of WED-flow on
business process management software design, by means of an easy to use, reliable and scalable tool that is capable to 
provide the flexibility demanded by modern transaction control systems. 
\\
\indent The remaining of this paper is structured as follows: Section II presents WED-SQL general software architecture aspects and
internal data structures. Section III briefly describes how the system works. A discussion about the algorithms involved
is presented in section IV. Following, a detailed description of transactional control and the communication protocol used
for external communication is given in section V. Finally, on section VI, we discuss the next challenges to improve and
expand this framework capabilities.
